---
title: "NatPatSIP - Martha's Rule Quarterly Presentation"
author: "Patient Safety Analysis Team"
format: 
   html:  
     embed-resources: true
     df-print: kable
     toc: true
     toc-location: left
     number-sections: true
execute:
  echo: false
  warning: false
params: 
  reporting_quarter: '2024/25 Q6'
  master_files_folder: '1. Master files'
  quarter_string: '25_26_Q1'
  marthas_data_file: 'marthas_psc_submissions_2025_26_Q1_processed_2025_09_08_184653.csv'
---


```{r}
# read in required packages
library(here)
library(tidyverse)
library(janitor)
library(lubridate)
library(ggplot2)
library(readxl)
library(forcats)
library(scales)
library(ggsci)
library(viridis)
library(Microsoft365R)
library(ggplot2) # likert plots
library(ggstats)# likert plots
library(kableExtra) # tables 
library(DT) # sortable tables

source('config_sharepoint_location.R')


# params
reporting_quarter <- params$reporting_quarter
master_files_folder <- params$master_files_folder
quarter_string <- params$quarter_string
marthas_data_file <- params$marthas_data_file

# bring site look up with postcode info and get geo inpho
site_postcode_list <- read.csv(here('lookups', 'psc_trust_site_lookup.csv')) |>
  select(trust_site_code, api_site_postcode) 

library(PostcodesioR)

postcodes_lookup_table <- lapply(unique(site_postcode_list$api_site_postcode), PostcodesioR::postcode_lookup) |>
  bind_rows() |>
  right_join(site_postcode_list, by = c('postcode' = 'api_site_postcode'))
```


```{r}
# NOTE block below is a place holder that willcoemi n place in 2526 Q2
# go into share point and download latest data
#master_files_dr <- chosenlib$get_item(glue::glue("{base_url}/{master_files_folder}"))

# marthas_data_file <- master_files_dr$list_files() |> 
#   select(name) |> 
#   filter(str_detect(name, 
#                     # below parameter is generated in read_qart_submissions.R
#                     quarter_string)
#          ) |>
#   as.vector() |>
#   unlist() |> 
#   unname()
# 
# marthas_identified_data <- master_files_dr$get_item(str_glue({marthas_data_file}))
# 
# marthas_identified_data$download(dest = here("data", str_glue({marthas_data_file})), 
#                        overwrite = T)

# marthas_data <- read.csv(here(str_glue('data/{marthas_data_file}'))) |>
#   mutate(updated_psc_name = str_remove_all(updated_psc_name, 'HIN'))

marthas_data_raw <- read.csv(here(str_glue('output/{marthas_data_file}'))) |>
  mutate(updated_psc_name = str_remove_all(updated_psc_name, ' HIN')) 

marthas_data_postcode  <- marthas_data_raw |>
  left_join(postcodes_lookup_table, by = c('site_code' = 'trust_site_code')) |>
  relocate(postcode, .after = name_of_site)
```

## Note

The visualisations and breakdowns below are experimental and for internal use only.

## Distribution of sites, by phase 

Missing data for two phase 2 sites here: 

- RGD03 - Leeds General Infirmary
- RAN01 - The Royal National Orthopaedic Hospital (Stanmore)

Are we missing one phase 2 site?

```{r}
# free styling aggregations and visualisations
marthas_data_to_aggregate <- marthas_data_raw |>
  rename('PSC' = updated_psc_name,
         'Phase' = phase)

marthas_data_to_aggregate |>
  tabyl(Phase, sort = T) |> 
  rename('Count' = n, 'Percent' = percent) |> 
  mutate(Percent = label_percent(accuracy = 0.1)(Percent)) |>
  kable(align = 'lll')
```



```{r}
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
# Get United Kingdom boundaries at 1:10m scale
uk <- ne_states(country = "United Kingdom", returnclass = "sf")

# Filter for England
england <- uk %>% filter(geonunit == "England")

# outilinee of england
england_outline <- england %>%
  summarise(geometry = st_union(geometry))

# make data sf
trust_sites_sf <- sf::st_as_sf(marthas_data_postcode , coords = c('longitude', 'latitude'),
                      crs = 4326)
```


```{r, fig.height=8,fig.width=8}
ggplot() +
  geom_sf(data = england_outline, fill = "white", color = "black") +
  geom_sf(data = trust_sites_sf, 
          aes(colour = factor(phase)),
          size = 2,
          alpha = 0.7
          ) +
  labs(title = "Martha's Rule sites in England, by phase")+
  scale_colour_manual(values = c('#003087', '#41B6E6'),
                      labels = c("Phase 1", "Phase 2")  # Custom legend labels
                    ) +
  theme_minimal()  +
  theme(axis.text = element_blank(),
        panel.grid = element_blank(),
        legend.title = element_blank()
        )
```

## Share of sites by PSC

::: {.panel-tabset}

### Table 

```{r}
# maybe make this into a bar plot, 2 shades of clue for each phase that make up the totality of sites 
share_of_sites <- marthas_data_to_aggregate |>
  tabyl(PSC, Phase) |>
  rename('Phase 1' = `1`,
         'Phase 2' = `2`
         ) 

share_of_sites_long <- share_of_sites |>
  pivot_longer(cols = starts_with('Phase'), 
               names_to = 'Phase', 
               values_to = 'Number of sites')

#share of site table 
share_of_sites_table <- share_of_sites |>
  adorn_totals('both') |>
  mutate(Percent = label_percent(accuracy = 0.1)(Total/210))

share_of_sites_table |> 
  kable(align = 'llll')

```

### Plot

```{r, fig.height=7,fig.width=10}
share_of_sites_for_plot <- share_of_sites_long |>
  left_join(share_of_sites_table |> select(PSC, Percent), by = 'PSC')

share_of_sites_perc_labels <- share_of_sites_for_plot |>
  group_by(PSC) |> 
  summarise('Number of sites' = sum(`Number of sites`), Percent = first(Percent)) |>
  ungroup()

ggplot(data = share_of_sites_for_plot,
       aes(x = `Number of sites`, y = forcats::fct_rev(PSC)
           )
       ) +
  geom_bar(aes( fill = forcats::fct_rev(Phase)),stat = 'identity', alpha = 0.7) +
  geom_text(data = share_of_sites_perc_labels,
            aes(x = `Number of sites`, 
                y = forcats::fct_rev(PSC), 
                label = Percent
                ),
            hjust = -0.1,
            position=position_dodge2(width=0.9),
            size = 4.5
            ) +
  scale_x_continuous(expand = expansion(mult = c(0.02,0.11))) +
  scale_fill_manual(values = c('#41B6E6', '#003087')) +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(color = "black", size = 13),
        axis.text.x = element_text(color = "black", size = 13),
        axis.text.y = element_text(color = "black", size = 13),
        panel.grid.minor.x = element_blank(),
        axis.ticks = element_line(color = "grey"),
        legend.title = element_blank())
```

:::

## Adult questions 

### Question 1

#### National overview

Lot of the data in the middle stages for either phase, even though sites in phase 1 are assumed to have been in the programme longer (check this assumption)

::: {.panel-tabset}

#### Plot

```{r, fig.height=7,fig.width=10}
national_by_phase <- marthas_data_to_aggregate |>
  mutate(Phase = as.character(Phase)) |>
  # exclude these from percetnage calcualtion
  filter(adults_patient_check_in != 'No adult patients'#,
         #Phase == '1'
         ) |>
  group_by(Phase, adults_patient_check_in) |>
  summarise(number_of_sites = n()) |>
  group_by(Phase) |>
  mutate(group_size_by_phase = sum(number_of_sites)) |>
  arrange(adults_patient_check_in)

ggplot(data = national_by_phase,
       aes(y = number_of_sites,
           x = factor(adults_patient_check_in)
           )
       ) +
  geom_bar(aes(fill = fct_rev(factor(Phase))),
           stat = 'identity',
           alpha = 0.7,
           position = position_dodge()) +
  labs(x = "Stage",
       y = "Number of sites",
       #fill = "Phase"
       title = "Stages of adoption of component 1 across Martha's rule sites, by phase",
       )+
  scale_fill_manual(values = c('#41B6E6', '#003087'),
                    name = "Phase",  # Custom legend title
                    labels = c("Phase 2", "Phase 1")  # Custom legend labels
                    )+
  scale_x_discrete(labels = label_wrap(25)) +
  theme_minimal() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_text(color = "black", size = 13),
        axis.text.x = element_text(color = "black", size = 13,
                                   angle = 90),
        axis.text.y = element_text(color = "black", size = 13),
        panel.grid.minor.x = element_blank(),
        axis.ticks = element_line(color = "grey"),
        legend.title = element_blank())
```

#### Table 

Note, fewer sites here (203) than overall, given there are sites with no adult units 
```{r}
national_all_phases <- marthas_data_to_aggregate |>
  # exclude these from percetnage calcualtion
  filter(adults_patient_check_in != 'No adult patients') |>
  tabyl(adults_patient_check_in) |>
  rename('Number of sites' = n, 'Percent' = percent, 'Stage' = adults_patient_check_in) |> 
  adorn_totals('row') |>
  mutate(Percent = label_percent(accuracy = 0.1)(Percent)) 

national_all_phases
```


#### Reports at or below stage 2
```{r}
marthas_data_to_aggregate |>
  mutate(Phase = as.character(Phase)) |>
  filter(adults_patient_check_in != 'No adult patients',
         str_detect(adults_patient_check_in, '(?i)stage (0|1|2)')) |>
  select(PSC, site_code, name_of_site, Phase, adults_patient_check_in) |>
  arrange(adults_patient_check_in, Phase, PSC) |>
  kable()
```

:::

#### PSC Overview

PSC breakdown of responses, by stage

::: {.panel-tabset}

#### Plot

```{r}
psc_all_phases <- marthas_data_to_aggregate |>
  # exclude these from percetnage calcualtion
  filter(adults_patient_check_in != 'No adult patients') |>
  select(PSC, site_code, adults_patient_check_in) |>
  mutate(stage_numeric = str_extract(adults_patient_check_in, '\\b[0-7]\\b'),
         stage_numeric = factor(stage_numeric))

#psc_all_phases

psc_all_phases_likert <- psc_all_phases |>
  select(-adults_patient_check_in) |>
  pivot_wider(names_from = PSC,
              values_from = stage_numeric)

#psc_all_phases_likert

gglikert(psc_all_phases_likert |> 
           # drop so we can graph by PSC correctly
           select(-site_code)
         )
```

#### Table
```{r}
psc_all_phases_table <- marthas_data_to_aggregate |>
  # exclude these from percetnage calcualtion
  filter(adults_patient_check_in != 'No adult patients') |>
  group_by(PSC, adults_patient_check_in, .drop = F) |>
  summarise(n_sites = n()) |>
  ungroup() |>
  group_by(PSC) |>
  mutate(total_n_psc_sites = sum(n_sites),
         Percent = n_sites/total_n_psc_sites,
         Percent = label_percent(accuracy = 0.1)(Percent)
         ) |>
  ungroup()

psc_all_phases_table |> 
  DT::datatable(options = list(pageLength = 10))
```

:::  

## All adult data 

<div style="width: 80vw;">

```{r}
# not sure this is helpful
stages <- c('Stage 0 - No contact', 'Stage 1 - Knowledge', 'Stage 2 - Interest')

marthas_data_to_aggregate |>
  select(PSC, site_code, name_of_site, Phase, matches('adults'))|>
  DT::datatable(options = list(pageLength = 21)) |>
  DT::formatStyle(
    columns = c('adults_patient_check_in', 'adults_independent_clinical_review',
                'adults_escalation_available_to_patient_carers'),
    backgroundColor = styleEqual(
      levels = stages,
      c('red', 'tomato', 'orange')
    )
  )
```

</div>

<div style="width: 75vw;">

## All paediatric data 
```{r}
marthas_data_to_aggregate |>
  select(PSC, site_code, name_of_site, Phase, matches('paediatric'))|>
  DT::datatable(options = list(pageLength = 21)) |>
  DT::formatStyle(
    columns = c('paediatric_patient_check_in', 'paediatric_independent_clinical_review',
                'paediatric_escalation_available_to_patient_carers'),
    backgroundColor = styleEqual(
      levels = stages,
      c('red', 'tomato', 'orange')
    )
  )
```

</div>
