---
title: "NatPatSIP - Martha's Rule QART Report"
author: "Patient Safety Analysis Team"
format: 
   html:  
     embed-resources: true
     df-print: kable
     toc: true
     toc-location: left
     number-sections: true
execute:
  echo: false
  warning: false
params: 
  reporting_quarter: '2025/26 Q1'
  master_files_folder: '1. Master files'
  quarter_string: '25_26_Q1'
  marthas_data_file: 'marthas_psc_submissions_2025_26_Q1_processed_2025_09_08_184653.csv'
---


```{r}
library(here)
library(tidyverse)
library(janitor)
library(lubridate)
library(ggplot2)
library(readxl)
library(forcats)
library(scales)
library(ggsci)
library(viridis)
library(Microsoft365R)
# likert plots
library(ggplot2) 
library(ggstats)
# tables 
library(kableExtra) 
# sortable tables
library(DT) 
# post code and mapping libraries
library(PostcodesioR)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)

source('config_sharepoint_location.R')

# params
reporting_quarter <- params$reporting_quarter
master_files_folder <- params$master_files_folder
quarter_string <- params$quarter_string
marthas_data_file <- params$marthas_data_file

# bring site look up with postcode info and get geo inpho
site_postcode_list <- read.csv(here('lookups', 'psc_trust_site_lookup.csv')) |>
  select(trust_site_code, api_site_postcode) 

postcodes_lookup_table <- lapply(unique(site_postcode_list$api_site_postcode),
                                 PostcodesioR::postcode_lookup) |>
  bind_rows() |>
  right_join(site_postcode_list, by = c('postcode' = 'api_site_postcode'))
```


```{r}
# NOTE block below is a place holder that will coem in place in 2526 Q2
# go into share point and download latest data
#master_files_dr <- chosenlib$get_item(glue::glue("{base_url}/{master_files_folder}"))

# marthas_data_file <- master_files_dr$list_files() |> 
#   select(name) |> 
#   filter(str_detect(name, 
#                     # below parameter is generated in read_qart_submissions.R
#                     quarter_string)
#          ) |>
#   as.vector() |>
#   unlist() |> 
#   unname()
# 
# marthas_identified_data <- master_files_dr$get_item(str_glue({marthas_data_file}))
# 
# marthas_identified_data$download(dest = here("data", str_glue({marthas_data_file})), 
#                        overwrite = T)

# marthas_data <- read.csv(here(str_glue('data/{marthas_data_file}'))) |>
#   mutate(updated_psc_name = str_remove_all(updated_psc_name, 'HIN'))

marthas_data_raw <- read.csv(here(str_glue('output/{marthas_data_file}'))) |>
  mutate(updated_psc_name = str_remove_all(updated_psc_name, ' HIN')) 

marthas_data_postcode  <- marthas_data_raw |>
  left_join(postcodes_lookup_table, by = c('site_code' = 'trust_site_code')) |>
  relocate(postcode, .after = name_of_site)
```

## Note

The visualisations and breakdowns below are experimental and for internal use only. Data presented was reported in `r reporting_quarter`.

## Distribution of sites by phase 

Missing data for two phase 2 sites here: 

- RGD03 - Leeds General Infirmary
- RAN01 - The Royal National Orthopaedic Hospital (Stanmore)

```{r}
# free styling aggregations and visualisations
marthas_data_to_aggregate <- marthas_data_raw |>
  rename('PSC' = updated_psc_name,
         'Phase' = phase) |>
  # consolidate org code and names for table printing
  mutate(name_trust_code = paste0(name_of_trust, ' (', trust_code, ')'),
         name_site_code = paste0(name_of_site, ' (', site_code, ')'),
         Phase = as.character(Phase),
         .before = Phase) |>
  select(-c(trust_code, name_of_trust, site_code, name_of_site))


marthas_data_to_aggregate |>
  tabyl(Phase, sort = T) |> 
  rename('Number of sites' = n, 'Percent' = percent) |> 
  mutate(Percent = label_percent(accuracy = 0.1)(Percent)) |>
  kable(align = 'lll')
```



```{r}
# get UK boundaries 
uk <- ne_states(country = "United Kingdom", returnclass = "sf")

# filter for England
england <- uk %>% filter(geonunit == "England")

# outline of england
england_outline <- england %>%
  summarise(geometry = st_union(geometry))

# make data sf
trust_sites_sf <- sf::st_as_sf(marthas_data_postcode , coords = c('longitude', 'latitude'),
                      crs = 4326)
```

**Figure :** Martha's Rule sites in England, by phase

```{r, fig.height=8,fig.width=8}
ggplot() +
  geom_sf(data = england_outline, fill = "white", color = "black") +
  geom_sf(data = trust_sites_sf, 
          aes(colour = factor(phase)),
          size = 2,
          alpha = 0.7
          ) +
  scale_colour_manual(values = c('#801650', '#28A197'),
                      labels = c("Phase 1", "Phase 2")  # custom legend labels
                    ) +
  theme_minimal()  +
  theme(axis.text = element_blank(),
        panel.grid = element_blank(),
        legend.title = element_blank()
        )
```

## Share of sites by PSC

::: {.panel-tabset}

```{r}
share_of_sites <- marthas_data_to_aggregate |>
  tabyl(PSC, Phase) |>
  rename('Phase 1' = `1`,
         'Phase 2' = `2`
         ) 

share_of_sites_long <- share_of_sites |>
  pivot_longer(cols = starts_with('Phase'), 
               names_to = 'Phase', 
               values_to = 'Number of sites')

#share of site table 
share_of_sites_table <- share_of_sites |>
  adorn_totals('both') |>
  mutate(Percent = label_percent(accuracy = 0.1)(Total/210))
```

### Plot

**Figure :** Breakdown of Martha's Rule sites by PSC and phase

```{r, fig.height=10,fig.width=10}
share_of_sites_for_plot <- share_of_sites_long |>
  left_join(share_of_sites_table |> select(PSC, Percent), by = 'PSC')

share_of_sites_perc_labels <- share_of_sites_for_plot |>
  group_by(PSC) |> 
  summarise('Number of sites' = sum(`Number of sites`), Percent = first(Percent)) |>
  ungroup()

ggplot(data = share_of_sites_for_plot,
       aes(x = `Number of sites`, y = forcats::fct_rev(PSC)
           )
       ) +
  geom_bar(aes( fill = forcats::fct_rev(Phase)),stat = 'identity', alpha = 0.7) +
  geom_text(data = share_of_sites_perc_labels,
            aes(x = `Number of sites`, 
                y = forcats::fct_rev(PSC), 
                label = Percent
                ),
            hjust = -0.1,
            position=position_dodge2(width=0.9),
            size = 4.5
            ) +
  scale_x_continuous(expand = expansion(mult = c(0.02,0.11))) +
  scale_fill_manual(values = c('#28A197', '#801650')
                    ) +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(color = "black", size = 13),
        axis.text.x = element_text(color = "black", size = 13),
        axis.text.y = element_text(color = "black", size = 13),
        legend.text = element_text(color = "black", size = 13),
        panel.grid.minor.x = element_blank(),
        axis.ticks = element_line(color = "grey"),
        legend.title = element_blank())
```

### Table

```{r}
share_of_sites_table |> 
  kable(align = 'llll')
```

:::

## Stages of adoption across adult units 

### Component 1

Patients will be asked, at least daily, about how they are feeling, and if they are getting better or worse, and this information will be acted on in a structured way.

#### National overview

::: {.panel-tabset}

#### Plot

**Figure :** Stages of adoption of component 1 across Martha's Rule sites with adult units, by phase

```{r}
# function to aggregate by phase and component
national_by_phase <- function(df, component, exclude){
  # quote variable
  component <- enquo(component)
  
  national_by_phase <- df |>
    # exclude these from percentage calculation
    filter(!!component != exclude
           ) |>
    group_by(Phase, !!component) |>
    summarise(number_of_sites = n()) |>
    group_by(Phase) |>
    mutate(group_size_by_phase = sum(number_of_sites)) |>
    arrange(!!component )
  
  national_by_phase
}

# function to plot aggregated table from national_by_phase()
plot_national_by_phase <- function(df, component){
  
  # quote variable
  component <- enquo(component)
  
  plot_national_by_phase <- ggplot(
    data = df,
    aes(
      y = number_of_sites,
      x = factor(!!component)
      )
    ) +
  geom_bar(aes(fill = fct_rev(factor(Phase))),
           stat = 'identity',
           alpha = 0.7,
           position = position_dodge()) +
  labs(x = "Stage",
       y = "Number of sites"
       )+
  scale_fill_manual(values = c('#28A197', '#801650'),
                    name = "Phase",  # custom legend title
                    labels = c("Phase 2", "Phase 1")  # custom legend labels
                    )+
  scale_x_discrete(labels = label_wrap(25)) +
  theme_minimal() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_text(color = "black", size = 13),
        axis.text.x = element_text(color = "black", size = 13,
                                   angle = 90),
        axis.text.y = element_text(color = "black", size = 13),
        panel.grid.minor.x = element_blank(),
        axis.ticks = element_line(color = "grey"),
        legend.title = element_blank(),
        legend.text = element_text(color = "black", size = 13)
        )
  
  plot_national_by_phase
}
```

```{r, fig.height=8,fig.width=10}
# deploy functions
national_by_phase_table <- national_by_phase(df = marthas_data_to_aggregate,
                                             component = adults_patient_check_in,
                                              exclude = 'No adult patients'
                                             )

plot_national_by_phase(df = national_by_phase_table, 
                       component = adults_patient_check_in
                         )
```

#### Table 

*Note: Fewer sites here (203) than overall, given sites with no adult units* 

```{r}
national_all_phases <- function(df, component, exclude){
  # quote
  component <- enquo(component)
  
  national_all_phases <- df |>
    # exclude these from percentage calculation
    filter(!!component != exclude) |>
    tabyl(!!component) |>
    rename('Number of sites' = n, 'Percent' = percent, 'Stage' = !!component) |> 
    adorn_totals('row') |>
    mutate(Percent = label_percent(accuracy = 0.1)(Percent)) 
  
  national_all_phases
}

national_all_phases(df = marthas_data_to_aggregate, 
                    component = adults_patient_check_in,
                    exclude = 'No adult patients'
                    ) |>
  kable()
```

#### Reports at or below stage 2
```{r}
below_stage_x <- function(df, component, exclude, stage_filter = '(?i)stage (0|1|2)'){
  
  component <- enquo(component)

  below_stage_x <- df |>
    filter(!!component != exclude,
           str_detect(!!component , stage_filter)) |>
    select(PSC, name_trust_code, name_site_code, Phase, !!component) |>
    rename('Trust'= name_trust_code,
          'Site' = name_site_code) |>
    arrange(!!component , Phase, PSC)
  
  below_stage_x
    
}

below_stage_x(df = marthas_data_to_aggregate, 
              component = adults_patient_check_in,
              exclude = 'No adult patients'
              ) |>
  rename('Component 1' = adults_patient_check_in) |>
  kable()
```

:::

#### PSC Overview

::: {.panel-tabset}

#### All sites

**Figure :** Adoption stages for component 1 across all sites with adult units, by PSC

```{r}
# this will allow colours to remain fixed by stage, even when we split vis by phase
fixed_colours <- setNames(
  object = c("#D73027", "#F46D43", "#FDAE61", "#FEE08B", "#D9EF8B", "#A6D96A", "#66BD63", "#1A9850"),
  nm = c('Stage 0 - No contact', 'Stage 1 - Knowledge', 'Stage 2 - Interest', 'Stage 3 - Decision',
         'Stage 4 - Testing or Implementation', 'Stage 5 - Adoption', 'Stage 6 - Spread', 
         'Stage 7 - Sustained')
  )

# graph each component as a likert scale
likert_component <- function(df, component, exclude, by_phase = NULL){
  
  component <- enquo(component)
  
  if (is.null(by_phase)) {
    psc_all_phases <- df |>
      select(PSC, name_site_code, !!component) |>
      # exclude these from percentage calculation
      filter(!!component != exclude) |>
      # note change from quoted component to local component
      mutate(component = factor(!!component))

    psc_for_likert <- psc_all_phases
    } else if (by_phase == 'Phase 1') {
      psc_phase_1 <- df |>
        # exclude these from percentage calculation
        filter(!!component != exclude,
               Phase == '1') |>
        select(PSC, name_site_code, !!component) |>
        # note change from quoted component to local component
        mutate(component = factor(!!component))
      
      psc_for_likert <- psc_phase_1
    } else if (by_phase == 'Phase 2') {
      psc_phase_2 <- df |>
        # exclude these from percentage calculation
        filter(!!component != exclude,
               Phase == '2') |>
        select(PSC, name_site_code, !!component) |>
        # note change from quoted component to local component
        mutate(component = factor(!!component))
      
      psc_for_likert <- psc_phase_2
      }
  
  psc_for_likert

  psc_all_phases_likert <- psc_for_likert |>
    select(-!!component) |>
    pivot_wider(names_from = PSC,
                values_from = component)

  plot <- gglikert(psc_all_phases_likert |>
             # drop so we can graph by PSC correctly
             select(-name_site_code),
             # exclude second set of percentages to the sides
             add_totals = F,
           ) +
    scale_fill_manual(
      values = fixed_colours
    ) +
    guides(fill = guide_legend(nrow = 3, byrow = T))+
    theme(axis.text.x = element_blank(),
          axis.text.y = element_text(color = "black", size = 13),
          legend.text = element_text(color = "black", size = 13)
            )

  plot
}
```


```{r, fig.height=10,fig.width=11}
likert_component(df = marthas_data_to_aggregate, 
                 component = adults_patient_check_in, 
                 exclude = 'No adult patients'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table <- function(df, component, exclude){
  component <- enquo(component)

  psc_all_phases_table <- df |>
    # exclude these from percentage calculation
    filter(!!component != exclude) |>
    group_by(PSC, !!component, .drop = F) |>
    summarise(`Number of sites` = n()) |>
    ungroup() |>
    group_by(PSC) |>
    mutate(`Sites in PSC` = sum(`Number of sites`),
           Percent = `Number of sites`/`Sites in PSC`,
           Percent = label_percent(accuracy = 0.1)(Percent)
           ) |>
    ungroup()

  psc_all_phases_table
}

psc_all_phases_table(df = marthas_data_to_aggregate,
                     component = adults_patient_check_in,
                     exclude = 'No adult patients'
                     ) |>
  rename('Component 1' = adults_patient_check_in) |>
  DT::datatable(options = list(pageLength = 10))
```

#### Phase 1

**Figure :** Adoption stages for component 1 across phase 1 sites with adult units, by PSC

```{r, fig.height=10,fig.width=11}
likert_component(df = marthas_data_to_aggregate, 
                 component = adults_patient_check_in, 
                 exclude = 'No adult patients',
                 by_phase = 'Phase 1'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '1'),
                     component = adults_patient_check_in,
                     exclude = 'No adult patients'
                     ) |>
  rename('Component 1' = adults_patient_check_in) |>
  DT::datatable(options = list(pageLength = 10))
```


#### Phase 2

**Figure :** Adoption stages for component 1 across phase 2 sites with adult units, by PSC

```{r, fig.height=10,fig.width=11}
likert_component(df = marthas_data_to_aggregate, 
                 component = adults_patient_check_in, 
                 exclude = 'No adult patients',
                 by_phase = 'Phase 2'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '2'),
                     component = adults_patient_check_in,
                     exclude = 'No adult patients'
                     ) |>
  rename('Component 1' = adults_patient_check_in) |>
  DT::datatable(options = list(pageLength = 10))
```

:::

### Component 2

All staff will be able, at any time, to ask for a review from a different team if they are concerned that a patient is deteriorating, and they are not being responded to.

#### National overview

::: {.panel-tabset}

#### Plot

**Figure :** Stages of adoption of component 2 across Martha’s Rule sites with adult units, by phase

```{r, fig.height=8,fig.width=10}
national_by_phase_table <- national_by_phase(df = marthas_data_to_aggregate,
                                             component = adults_independent_clinical_review,
                                             exclude = 'No adult patients'
)

plot_national_by_phase(df = national_by_phase_table, 
                       component = adults_independent_clinical_review
)
```


#### Table 

```{r}
national_all_phases(df = marthas_data_to_aggregate, 
                    component = adults_independent_clinical_review,
                    exclude = 'No adult patients'
) |>
  kable()
```


#### Reports at or below stage 2

```{r}
below_stage_x(df = marthas_data_to_aggregate, 
              component = adults_independent_clinical_review,
              exclude = 'No adult patients'
              ) |>
  rename('Component 2' = adults_independent_clinical_review) |>
  kable()
```

:::

#### PSC Overview

::: {.panel-tabset}

#### All sites

**Figure :** Adoption stages for component 2 across all sites with adult units, by PSC

```{r, fig.height=10,fig.width=11}
likert_component(df = marthas_data_to_aggregate, 
                 component = adults_independent_clinical_review,
                 exclude = 'No adult patients'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate,
                     component = adults_independent_clinical_review,
                     exclude = 'No adult patients'
                     ) |>
  rename('Component 2' = adults_independent_clinical_review) |>
  DT::datatable(options = list(pageLength = 10))
```


#### Phase 1

**Figure :** Adoption stages for component 2 across phase 1 sites with adult units, by PSC

```{r, fig.height=10,fig.width=11}
likert_component(df = marthas_data_to_aggregate, 
                 component = adults_independent_clinical_review, 
                 exclude = 'No adult patients',
                 by_phase = 'Phase 1'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '1'),
                     component = adults_independent_clinical_review,
                     exclude = 'No adult patients'
                     ) |>
  rename('Component 2' = adults_independent_clinical_review) |>
  DT::datatable(options = list(pageLength = 10))
```


#### Phase 2

**Figure :** Adoption stages for component 2 across phase 2 sites with adult units, by PSC

```{r, fig.height=10,fig.width=11}
likert_component(df = marthas_data_to_aggregate, 
                 component = adults_independent_clinical_review, 
                 exclude = 'No adult patients',
                 by_phase = 'Phase 2'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '2'),
                     component = adults_independent_clinical_review,
                     exclude = 'No adult patients'
                     ) |>
  rename('Component 2' = adults_independent_clinical_review) |>
  DT::datatable(options = list(pageLength = 10))
```

:::

### Component 3

This escalation route will also always be available to patients themselves, their families and carers and advertised across the hospital.

#### National overview

::: {.panel-tabset}

#### Plot

**Figure :** Stages of adoption of component 3 across Martha’s Rule sites with adult units, by phase

```{r, fig.height=8,fig.width=10}
national_by_phase_table <- national_by_phase(df = marthas_data_to_aggregate,
                                             component = adults_escalation_available_to_patient_carers,
                                             exclude = 'No adult patients'
)

plot_national_by_phase(df = national_by_phase_table, 
                       component = adults_escalation_available_to_patient_carers
)
```


#### Table 

```{r}
national_all_phases(df = marthas_data_to_aggregate, 
                    component = adults_escalation_available_to_patient_carers,
                    exclude = 'No adult patients'
) |>
  kable()
```


#### Reports at or below stage 2

```{r}
below_stage_x(df = marthas_data_to_aggregate, 
              component = adults_escalation_available_to_patient_carers,
              exclude = 'No adult patients'
              ) |>
  rename('Component 3' = adults_escalation_available_to_patient_carers) |>
  kable()
```

:::

#### PSC Overview

::: {.panel-tabset}

#### All Sites

**Figure :** Adoption stages for component 3 across all sites with adult units, by PSC

```{r, fig.height=10,fig.width=11}
likert_component(df = marthas_data_to_aggregate, 
                 component = adults_escalation_available_to_patient_carers,
                 exclude = 'No adult patients'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate,
                     component = adults_escalation_available_to_patient_carers,
                     exclude = 'No adult patients'
                     ) |>
  rename('Component 3' = adults_escalation_available_to_patient_carers) |>
  DT::datatable(options = list(pageLength = 10))
```


#### Phase 1

**Figure :** Adoption stages for component 3 across phase 1 sites with adult units, by PSC

```{r, fig.height=10,fig.width=11}
likert_component(df = marthas_data_to_aggregate, 
                 component = adults_escalation_available_to_patient_carers, 
                 exclude = 'No adult patients',
                 by_phase = 'Phase 1'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '1'),
                     component = adults_escalation_available_to_patient_carers,
                     exclude = 'No adult patients'
                     ) |>
  rename('Component 3' = adults_escalation_available_to_patient_carers) |>
  DT::datatable(options = list(pageLength = 10))
```


#### Phase 2

**Figure :** Adoption stages for component 3 across phase 2 sites with adult units, by PSC

```{r, fig.height=10,fig.width=11}
likert_component(df = marthas_data_to_aggregate, 
                 component = adults_escalation_available_to_patient_carers, 
                 exclude = 'No adult patients',
                 by_phase = 'Phase 2'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '2'),
                     component = adults_escalation_available_to_patient_carers,
                     exclude = 'No adult patients'
                     ) |>
  rename('Component 3' = adults_escalation_available_to_patient_carers) |>
  DT::datatable(options = list(pageLength = 10))
```

:::

## Stages of adoption across paediatric units 

### Component 1

Patients will be asked, at least daily, about how they are feeling, and if they are getting better or worse, and this information will be acted on in a structured way.

#### National overview

::: {.panel-tabset}

#### Plot

**Figure :** Stages of adoption of component 1 across Martha’s Rule sites with paediatric units, by phase

```{r, fig.height=8,fig.width=10}
national_by_phase_table <- national_by_phase(df = marthas_data_to_aggregate,
                                             component = paediatric_patient_check_in,
                                              exclude = 'No paediatric patients'
)

plot_national_by_phase(df = national_by_phase_table, 
                       component = paediatric_patient_check_in
)
```


#### Table 

*Note: fewer sites here (180) than overall, given sites with no paediatric units*

```{r}
national_all_phases(df = marthas_data_to_aggregate, 
                    component = paediatric_patient_check_in,
                    exclude = 'No paediatric patients'
) |>
  kable()
```


#### Reports at or below stage 2

```{r}
below_stage_x(df = marthas_data_to_aggregate, 
              component = paediatric_patient_check_in,
              exclude = 'No paediatric patients'
              ) |>
  rename('Component 1' = paediatric_patient_check_in) |>
  kable()
```

:::

#### PSC Overview

::: {.panel-tabset}

#### All sites

**Figure :** Adoption stages for component 1 across all sites with paediatric units, by PSC

```{r, fig.height=10,fig.width=11}
likert_component(df = marthas_data_to_aggregate, 
                 component = paediatric_patient_check_in, 
                 exclude = 'No paediatric patients'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate,
                     component = paediatric_patient_check_in,
                     exclude = 'No paediatric patients'
                     ) |>
  rename('Component 1' = paediatric_patient_check_in) |>
  DT::datatable(options = list(pageLength = 10))
```

#### Phase 1

**Figure :** Adoption stages for component 1 across phase 1 sites with paediatric units, by PSC

```{r, fig.height=10,fig.width=11}
likert_component(df = marthas_data_to_aggregate, 
                 component = paediatric_patient_check_in, 
                 exclude = 'No paediatric patients',
                 by_phase = 'Phase 1'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '1'),
                     component = paediatric_patient_check_in,
                     exclude = 'No paediatric patients'
                     ) |>
  rename('Component 1' = paediatric_patient_check_in) |>
  DT::datatable(options = list(pageLength = 10))
```


#### Phase 2

**Figure :** Adoption stages for component 1 across phase 2 sites with paediatric units, by PSC

```{r, fig.height=10,fig.width=11}
likert_component(df = marthas_data_to_aggregate, 
                 component = paediatric_patient_check_in, 
                 exclude = 'No paediatric patients',
                 by_phase = 'Phase 2'
                   )

```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '2'),
                     component = paediatric_patient_check_in,
                     exclude = 'No paediatric patients'
                     ) |>
  rename('Component 1' = paediatric_patient_check_in) |>
  DT::datatable(options = list(pageLength = 10))
```

:::

### Component 2

All staff will be able, at any time, to ask for a review from a different team if they are concerned that a patient is deteriorating, and they are not being responded to.

#### National overview

::: {.panel-tabset}

#### Plot

**Figure :** Stages of adoption of component 2 across Martha’s Rule sites with paediatric units, by phase

```{r, fig.height=8,fig.width=10}
national_by_phase_table <- national_by_phase(df = marthas_data_to_aggregate,
                                             component = paediatric_independent_clinical_review,
                                              exclude = 'No paediatric patients'
)

plot_national_by_phase(df = national_by_phase_table, 
                       component = paediatric_independent_clinical_review
)
```


#### Table 

```{r}
national_all_phases(df = marthas_data_to_aggregate, 
                    component = paediatric_independent_clinical_review,
                    exclude = 'No paediatric patients'
) |>
  kable()
```


#### Reports at or below stage 2

```{r}
below_stage_x(df = marthas_data_to_aggregate, 
              component = paediatric_independent_clinical_review,
              exclude = 'No paediatric patients'
              ) |>
  rename('Component 2' = paediatric_independent_clinical_review) |>
  kable()
```

:::

#### PSC Overview

::: {.panel-tabset}

#### All sites

**Figure :** Adoption stages for component 2 across all sites with paediatric units, by PSC

```{r, fig.height=10,fig.width=11}
likert_component(df = marthas_data_to_aggregate, 
                 component = paediatric_independent_clinical_review, 
                 exclude = 'No paediatric patients'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate,
                     component = paediatric_independent_clinical_review,
                     exclude = 'No paediatric patients'
                     ) |>
  rename('Component 2' = paediatric_independent_clinical_review) |>
  DT::datatable(options = list(pageLength = 10))
```

#### Phase 1

**Figure :** Adoption stages for component 2 across phase 1 sites with paediatric units, by PSC

```{r, fig.height=10,fig.width=11}
likert_component(df = marthas_data_to_aggregate, 
                 component = paediatric_independent_clinical_review, 
                 exclude = 'No paediatric patients',
                 by_phase = 'Phase 1'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '1'),
                     component = paediatric_independent_clinical_review,
                     exclude = 'No paediatric patients'
                     ) |>
  rename('Component 2' = paediatric_independent_clinical_review) |>
  DT::datatable(options = list(pageLength = 10))
```


#### Phase 2

**Figure :** Adoption stages for component 2 across phase 2 sites with paediatric units, by PSC

```{r, fig.height=10,fig.width=11}
likert_component(df = marthas_data_to_aggregate, 
                 component = paediatric_independent_clinical_review, 
                 exclude = 'No paediatric patients',
                 by_phase = 'Phase 2'
                   )

```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '2'),
                     component = paediatric_independent_clinical_review,
                     exclude = 'No paediatric patients'
                     ) |>
  rename('Component 2' = paediatric_independent_clinical_review) |>
  DT::datatable(options = list(pageLength = 10))
```

:::

### Component 3

This escalation route will also always be available to patients themselves, their families and carers and advertised across the hospital.

#### National overview

::: {.panel-tabset}

#### Plot

**Figure :** Stages of adoption of component 3 across Martha’s Rule sites with paediatric units, by phase

```{r, fig.height=8,fig.width=10}
national_by_phase_table <- national_by_phase(df = marthas_data_to_aggregate,
                                             component = paediatric_escalation_available_to_patient_carers,
                                              exclude = 'No paediatric patients'
)

plot_national_by_phase(df = national_by_phase_table, 
                       component = paediatric_escalation_available_to_patient_carers
)
```


#### Table 

```{r}
national_all_phases(df = marthas_data_to_aggregate, 
                    component = paediatric_escalation_available_to_patient_carers,
                    exclude = 'No paediatric patients'
) |>
  kable()
```


#### Reports at or below stage 2

```{r}
below_stage_x(df = marthas_data_to_aggregate, 
              component = paediatric_escalation_available_to_patient_carers,
              exclude = 'No paediatric patients'
              ) |>
  rename('Component 3' = paediatric_escalation_available_to_patient_carers) |>
  kable()
```

:::

#### PSC Overview

::: {.panel-tabset}

#### All sites

**Figure :** Adoption stages for component 3 across all sites with paediatric units, by PSC

```{r, fig.height=10,fig.width=11}
likert_component(df = marthas_data_to_aggregate, 
                 component = paediatric_escalation_available_to_patient_carers, 
                 exclude = 'No paediatric patients'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate,
                     component = paediatric_escalation_available_to_patient_carers,
                     exclude = 'No paediatric patients'
                     ) |>
  rename('Component 3' = paediatric_escalation_available_to_patient_carers) |>
  DT::datatable(options = list(pageLength = 10))
```

#### Phase 1

**Figure :** Adoption stages for component 3 across phase 1 sites with paediatric units, by PSC

```{r, fig.height=10,fig.width=11}
likert_component(df = marthas_data_to_aggregate, 
                 component = paediatric_escalation_available_to_patient_carers, 
                 exclude = 'No paediatric patients',
                 by_phase = 'Phase 1'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '1'),
                     component = paediatric_escalation_available_to_patient_carers,
                     exclude = 'No paediatric patients'
                     ) |>
  rename('Component 3' = paediatric_escalation_available_to_patient_carers) |>
  DT::datatable(options = list(pageLength = 10))
```


#### Phase 2

**Figure :** Adoption stages for component 3 across phase 2 sites with paediatric units, by PSC
```{r, fig.height=10,fig.width=11}
likert_component(df = marthas_data_to_aggregate, 
                 component = paediatric_escalation_available_to_patient_carers, 
                 exclude = 'No paediatric patients',
                 by_phase = 'Phase 2'
                   )

```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '2'),
                     component = paediatric_escalation_available_to_patient_carers,
                     exclude = 'No paediatric patients'
                     ) |>
  rename('Component 3' = paediatric_escalation_available_to_patient_carers) |>
  DT::datatable(options = list(pageLength = 10))
```

:::

## All adult data 

<div style="width: 75vw;">

```{r}
# stages to colour
stages <- c('Stage 0 - No contact', 'Stage 1 - Knowledge', 'Stage 2 - Interest')

marthas_data_to_aggregate |>
  select(PSC, name_trust_code, name_site_code, 
         Phase, matches('adults')) |>
  rename('Trust' = name_trust_code, 
         'Site' = name_site_code,
         'Component 1' = adults_patient_check_in,
         'Component 2' = adults_independent_clinical_review,
         'Component 3' = adults_escalation_available_to_patient_carers
         ) |>
    DT::datatable(options = list(pageLength = 21)) |>
    DT::formatStyle(
      columns = c('Component 1', 'Component 2','Component 3'),
      backgroundColor = styleEqual(
        levels = stages,
        c('red', 'tomato', 'orange')
      )
    ) 
```

</div>

<div style="width: 75vw;">

## All paediatric data 
```{r}
marthas_data_to_aggregate |>
  select(PSC, name_trust_code, name_site_code, Phase, matches('paediatric')) |> 
  rename('Trust' = name_trust_code, 
         'Site' = name_site_code,
         'Component 1' = paediatric_patient_check_in,
         'Component 2' = paediatric_independent_clinical_review,
         'Component 3' = paediatric_escalation_available_to_patient_carers
         ) |>
  DT::datatable(options = list(pageLength = 21)) |>
  DT::formatStyle(
    columns = c('Component 1', 'Component 2','Component 3'),
    backgroundColor = styleEqual(
      levels = stages,
      c('red', 'tomato', 'orange')
    )
  )
```

</div>
