---
title: "NatPatSIP - Martha's Rule QART Report"
author: "Patient Safety Analysis Team"
format: 
   html:  
     embed-resources: true
     df-print: kable
     toc: true
     toc-location: left
     number-sections: true
execute:
  echo: false
  warning: false
params: 
  reporting_quarter: '2026/27 Q1'
  master_files_folder: '1. Master files'
  quarter_string: '2026_27_Q1'
---

```{r}
library(here)
library(tidyverse)
library(janitor)
library(lubridate)
library(ggplot2)
library(readxl)
library(forcats)
library(scales)
library(ggsci)
library(viridis)
library(Microsoft365R)
# likert plots
library(ggplot2) 
library(ggstats)
# tables 
library(kableExtra) 
# sortable tables
library(DT) 
# post code and mapping libraries
library(PostcodesioR)
library(sf)

source('config_sharepoint_location.R')

# params
reporting_quarter <- params$reporting_quarter
master_files_folder <- params$master_files_folder
quarter_string <- params$quarter_string
```

```{r}
# go into share point and download lookup with postcode info for map
master_files_dr <- chosenlib$get_item(glue::glue(
  "{base_url}/{master_files_folder}"))

lookup_data_file <- master_files_dr$list_files() |>
  select(name) |>
  filter(str_detect(name,'psc_trust_site_lookup.csv'
                    )
         ) |>
  as.vector() |>
  unlist() |>
  unname()

lookup_file_identified <- master_files_dr$get_item(str_glue({lookup_data_file}))

lookup_file_identified$download(dest = here("data", str_glue({lookup_data_file})),
                       overwrite = T)

# bring site look up with postcode info and get geo info
site_postcode_list <- read.csv(here('data', 'psc_trust_site_lookup.csv')) |>
  select(site_sdcs_code, api_site_postcode)

postcodes_lookup_table <- lapply(unique(site_postcode_list$api_site_postcode),
                                 PostcodesioR::postcode_lookup) |>
  bind_rows() |>
  right_join(site_postcode_list, by = c('postcode' = 'api_site_postcode'))

# go into share point and download latest data
marthas_data_file <- master_files_dr$list_files() |>
  select(name) |>
  filter(str_detect(name,
                    # below parameter is generated in read_qart_submissions.R
                    paste0('marthas_qart_adults_paeds_data_upto_', quarter_string)
                    )
         ) |>
  as.vector() |>
  unlist() |>
  unname()

marthas_identified_data <- master_files_dr$get_item(str_glue({marthas_data_file}))

marthas_identified_data$download(dest = here("data", str_glue({marthas_data_file})),
                       overwrite = T)


marthas_data <- read.csv(here(str_glue('data/{marthas_data_file}'))) |>
  mutate(updated_psc_name = str_remove_all(updated_psc_name, 'HIN'))

latest_quarter <- max(marthas_data$quarter)

marthas_data_raw <- marthas_data |>
  filter(quarter == latest_quarter)

marthas_data_postcode  <- marthas_data_raw |>
  left_join(postcodes_lookup_table, by = c('site_code' = 'site_sdcs_code')) |>
  relocate(postcode, .after = name_of_site)
```

## User guidance notes

- The visualisations and breakdowns below are experimental and for NHS England's internal use only. Data presented was reported in `r reporting_quarter`.
- Several breakdowns of the data are presented throughout the report. **Always check the size of numerators and denominators behind a percentage.** The more granular a breakdown, the higher the chances of obtaining 0% and 100% values due to small group sizes.
- PSC overview charts for every component have been "centred" to allow comparisons of adoption maturity between PSC's. This means the middle of the chart is a focal point that represents the most common stage reported. Bars will sway to the right or left of that focal point to reflect adoption progress. Bars leaning towards the right-hand side of the chart suggest more mature adoption, whilst bars leaning to the left suggest less mature adoption.
- To get the best experience out of this interactive report, **please make sure you are opening a local html file copy of this file.** Do not open this report directly on SharePoint. Instead, hover over the file name and click the `...` icon to download a copy of the file. Then open from your downloads folder, which should display the file in your browser.

## Distribution of sites by phase

```{r}
# aggregations and visualisations
marthas_data_to_aggregate <- marthas_data_raw |>
  rename('PSC' = updated_psc_name,
         'Phase' = phase) |>
  mutate(
    # fix trust names
    name_of_trust = str_to_title(name_of_trust), # turn into title case
    # then fix strings: Nhs, And, Of
    name_of_trust = str_replace_all(name_of_trust, '\\bNhs\\b', str_to_upper),
    name_of_trust = str_replace_all(name_of_trust, '\\bAnd\\b|\\bOf\\b', str_to_lower),
    # repeat for site names
    name_of_site = str_to_title(name_of_site),
    name_of_site = str_replace_all(name_of_site, '\\bNhs|Hq\\b', str_to_upper),
    name_of_site = str_replace_all(name_of_site, '\\bAnd\\b|\\bOf\\b', str_to_lower),
    # consolidate org code and names for table printing
    name_trust_code = paste0(name_of_trust, ' (', trust_code, ')'),
    name_site_code = paste0(name_of_site, ' (', site_code, ')'),
    Phase = as.character(Phase),
    .before = Phase) |>
  select(-c(trust_code, name_of_trust, site_code, name_of_site))


marthas_data_to_aggregate |>
  tabyl(Phase, sort = T) |> 
  rename('Number of sites' = n, 'Percent' = percent) |> 
  mutate(Percent = label_percent(accuracy = 0.1)(Percent)) |>
  kable(align = 'lll')
```

```{r}
# get UK boundaries
uk <- st_read(
  "https://open-geography-portalx-ons.hub.arcgis.com/datasets/ons::countries-december-2025-boundaries-uk-buc.geojson",
  #suppress output message
  quiet = T
)

# filter for England
england <- uk[uk$CTRY25NM == "England", ]

# outline of england
england_outline <- england |>
  summarise(geometry = st_union(geometry))

# make data sf
trust_sites_sf <- sf::st_as_sf(marthas_data_postcode,
                               coords = c('longitude','latitude'),
                               crs = 4326)
```

**Figure 1:** Martha's Rule sites in England, by phase

```{r, fig.height=8,fig.width=8.2}
ggplot() +
  geom_sf(data = england_outline, fill = "white", color = "black") +
  geom_sf(data = trust_sites_sf,
          aes(colour = factor(phase)),
          size = 2
          ) +
  scale_colour_manual(values = c('#801650', '#28A197'),
                      labels = c("Phase 1", "Phase 2")  # custom legend labels
                    ) +
  theme_minimal()  +
  theme(axis.text = element_blank(),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        legend.position = 'top',
        legend.justification = 'left',
        legend.text = element_text(size = 14)
        )
```

## Share of sites by PSC

::: panel-tabset
```{r}
share_of_sites <- marthas_data_to_aggregate |>
  tabyl(PSC, Phase) |>
  rename('Phase 1' = `1`,
         'Phase 2' = `2`
         ) 

share_of_sites_long <- share_of_sites |>
  pivot_longer(cols = starts_with('Phase'), 
               names_to = 'Phase', 
               values_to = 'Number of sites')

#share of site table 
total_sites <- sum(share_of_sites_long$`Number of sites`)

share_of_sites_table <- share_of_sites |>
  adorn_totals('both') |>
  mutate(Percent = label_percent(accuracy = 0.1)(Total/total_sites))
```

### Plot

**Figure 2:** Breakdown of Martha's Rule sites by PSC and phase

```{r, fig.height=10,fig.width=10}
share_of_sites_for_plot <- share_of_sites_long |>
  left_join(share_of_sites_table |> select(PSC, Percent), by = 'PSC')

ggplot(data = share_of_sites_for_plot,
       aes(x = `Number of sites`, y = forcats::fct_rev(PSC)
           )
       ) +
  geom_bar(aes( fill = forcats::fct_rev(Phase)),stat = 'identity') +
  # label 
  geom_text(data = share_of_sites_table |> filter(PSC != 'Total'),
            aes(x = Total,
                y = forcats::fct_rev(PSC),
                label = Percent
            ),
            hjust = -0.1,
            position=position_dodge2(width=0.9),
            size = 4.5
            ) +
  scale_x_continuous(expand = expansion(mult = c(0.02,0.11))) +
  scale_fill_manual(values = c('#801650', '#28A197'),
                    breaks = c('Phase 1', 'Phase 2')
                    ) +
  theme_minimal() +
  theme(axis.title = element_blank(),
        axis.text.x = element_text(color = "black", size = 14),
        axis.text.y = element_text(color = "black", size = 14),
        legend.text = element_text(color = "black", size = 14),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        axis.ticks = element_line(color = "grey"),
        legend.title = element_blank(),
        legend.position = 'top',
        legend.justification = 'left',
        legend.location = 'plot'
        )
```

### Table

```{r}
share_of_sites_table |> 
  kable(align = 'llll')
```
:::

## Stages of adoption across adult units

```{r}
# function to aggregate spefic component by stage
national_all_phases <- function(df, component, exclude){
  # quote
  component <- enquo(component)
  
  national_all_phases <- df |>
    # exclude these from percentage calculation
    filter(!!component != exclude) |>
    tabyl(!!component) |>
    rename('Number of sites' = n, 'Percent' = percent, 'Stage' = !!component) 

  # stage check to show all stages in  tables
  all_possible_stages <- c('Stage 0 - No contact',
                           'Stage 1 - Knowledge',
                           'Stage 2 - Interest',
                           'Stage 3 - Decision',
                           'Stage 4 - Testing or Implementation',
                           'Stage 5 - Adoption',
                           'Stage 6 - Spread',
                           'Stage 7 - Sustained')
  
  # this is a tibble that has the stages not shown in aggregation by tabyl
  stages_zero_values <- tibble(Stage = all_possible_stages,
                              `Number of sites` = 0, 
                               Percent = 0) |>
    # we only keep what's not shown by tabyl
    filter(!Stage %in% national_all_phases$Stage)

  national_all_phases_tidy <- national_all_phases |>
    # add rows with zero values
    bind_rows(stages_zero_values)|>
    arrange(Stage) |>
    # add totals and format percentages
    adorn_totals('row') |>
    mutate(Percent = label_percent(accuracy = 0.1)(Percent))
    
  
  national_all_phases_tidy
}

adults_comp_1 <- national_all_phases(df = marthas_data_to_aggregate, 
                    component = adults_patient_check_in,
                    exclude = 'No adult patients'
                    )

total_sites_adult_units <- adults_comp_1 |> 
  filter(Stage == 'Total') |> 
  pull(`Number of sites`)
```

**Note:** The graphs and tables presented below include only sites that have adult units. Therefore, the total presented for this section (n = `r total_sites_adult_units`) is smaller than the total number of all Martha's Rule sites (n = `r total_sites`)

### Component 1

Patients will be asked, at least daily, about how they are feeling, and if they are getting better or worse, and this information will be acted on in a structured way.

#### National overview

::: panel-tabset
#### Plot

**Figure 3:** Stages of adoption of component 1 across Martha's Rule sites with adult units, by phase

```{r}
# function to aggregate by phase and component
national_by_phase <- function(df, component, exclude){
  # quote variable
  component <- enquo(component)
  
  national_by_phase <- df |>
    # exclude these from percentage calculation
    filter(!!component != exclude
           ) |>
    group_by(Phase, !!component) |>
    summarise(number_of_sites = n()) |>
    ungroup() |>
    group_by(Phase) |>
    mutate(group_size_by_phase = sum(number_of_sites)) |>
    ungroup() |>
    arrange(!!component )
  
  national_by_phase
}

# function to plot aggregated table from national_by_phase()
plot_national_by_phase <- function(df, component){
  
  # quote variable
  component <- enquo(component)
  
  plot_national_by_phase <- ggplot(
    data = df,
    aes(
      x = number_of_sites,
      y = factor(!!component)
      )
    ) +
  geom_bar(aes(fill = #fct_rev(factor(Phase))),
                 factor(Phase)),
           stat = 'identity',
           # preserve width of bars, even when there's no data for one group
           position = position_dodge(preserve = 'single')) +
  scale_fill_manual(values = c('#801650', '#28A197'),
                    labels = c('Phase 1', 'Phase 2')  # custom legend labels
                    )+
  scale_y_discrete(labels = label_wrap(25)) +
  theme_minimal() +
  theme(axis.title = element_blank(),
        axis.text = element_text(color = "black", size = 14),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        axis.ticks = element_line(color = "grey"),
        legend.title = element_blank(),
        legend.text = element_text(color = "black", size = 14),
        legend.position = 'top',
        legend.justification = 'left',
        legend.location = 'plot'
        )
  
  plot_national_by_phase
}
```

```{r, fig.height=8,fig.width=10}
# deploy functions
national_by_phase_table <- national_by_phase(df = marthas_data_to_aggregate,
                                             component = adults_patient_check_in,
                                              exclude = 'No adult patients'
                                             )

plot_national_by_phase(df = national_by_phase_table, 
                       component = adults_patient_check_in
                         )
```

#### Table

**All sites**

```{r}
adults_comp_1 |>
  kable() 
```

**Phase 1 sites**

```{r}
national_all_phases(df = marthas_data_to_aggregate |> 
                      filter(Phase == 1), 
                    component = adults_patient_check_in,
                    exclude = 'No adult patients'
                    ) |>
  kable()
```

**Phase 2 sites**

```{r}
national_all_phases(df = marthas_data_to_aggregate |> 
                      filter(Phase == 2), 
                    component = adults_patient_check_in,
                    exclude = 'No adult patients'
                    ) |>
  kable()
```

#### Reports at or below stage 3

```{r}
below_stage_x <- function(df, component, exclude, stage_filter = '(?i)stage (0|1|2|3)'){
  
  component <- enquo(component)

  below_stage_x <- df |>
    filter(!!component != exclude,
           str_detect(!!component , stage_filter)) |>
    select(PSC, name_trust_code, name_site_code, Phase, !!component) |>
    rename('Trust'= name_trust_code,
          'Site' = name_site_code) |>
    arrange(!!component , Phase, PSC)
  
  below_stage_x
    
}

below_stage_x(df = marthas_data_to_aggregate, 
              component = adults_patient_check_in,
              exclude = 'No adult patients'
              ) |>
  rename('Component 1' = adults_patient_check_in) |>
  kable()
```
:::

#### PSC Overview

::: panel-tabset
#### All sites

**Figure 4:** Adoption stages for component 1 across all sites with adult units, by PSC

```{r}
# this will allow colours to remain fixed by stage, even when we split vis by phase
fixed_colours <- setNames(
  object = c("#D73027", "#F46D43", "#FDAE61", "#FEE08B", "#D9EF8B", "#A6D96A", "#66BD63", "#1A9850"),
  nm = c('Stage 0 - No contact', 'Stage 1 - Knowledge', 'Stage 2 - Interest', 'Stage 3 - Decision',
         'Stage 4 - Testing or Implementation', 'Stage 5 - Adoption', 'Stage 6 - Spread', 
         'Stage 7 - Sustained')
  )

# graph each component as a likert scale
likert_component <- function(df, component, exclude, by_phase = NULL){
  
  component <- enquo(component)
  
  if (is.null(by_phase)) {
    psc_all_phases <- df |>
      select(PSC, name_site_code, !!component) |>
      # exclude these from percentage calculation
      filter(!!component != exclude) |>
      # note change from quoted component to local component
      mutate(component = factor(!!component))

    psc_for_likert <- psc_all_phases
    } else if (by_phase == 'Phase 1') {
      psc_phase_1 <- df |>
        # exclude these from percentage calculation
        filter(!!component != exclude,
               Phase == '1') |>
        select(PSC, name_site_code, !!component) |>
        # note change from quoted component to local component
        mutate(component = factor(!!component))
      
      psc_for_likert <- psc_phase_1
    } else if (by_phase == 'Phase 2') {
      psc_phase_2 <- df |>
        # exclude these from percentage calculation
        filter(!!component != exclude,
               Phase == '2') |>
        select(PSC, name_site_code, !!component) |>
        # note change from quoted component to local component
        mutate(component = factor(!!component))
      
      psc_for_likert <- psc_phase_2
      }
  
  psc_for_likert

  psc_all_phases_likert <- psc_for_likert |>
    select(-!!component) |>
    pivot_wider(names_from = PSC,
                values_from = component)

  plot <- gglikert(psc_all_phases_likert |>
             # drop so we can graph by PSC correctly
             select(-name_site_code),
             # exclude second set of percentages to the sides
             add_totals = F,
             labels_size = 4.5
           ) +
    scale_fill_manual(
      values = fixed_colours
    ) +
    guides(fill = guide_legend(nrow = 2, byrow = T))+
    theme(axis.text.x = element_blank(),
          axis.text.y = element_text(color = "black", size = 14),
          legend.text = element_text(color = "black", size = 14),
          legend.position = 'top',
          legend.justification = 'left',
          legend.location = 'plot'
          )

  plot
}
```

```{r, fig.height=10,fig.width=12}
likert_component(df = marthas_data_to_aggregate, 
                 component = adults_patient_check_in, 
                 exclude = 'No adult patients'
                   ) 
```

**Underlying data**

```{r}
psc_all_phases_table <- function(df, component, exclude){
  component <- enquo(component)

  psc_all_phases_table <- df |>
    # exclude these from percentage calculation
    filter(!!component != exclude) |>
    group_by(PSC, !!component, .drop = F) |>
    summarise(`Number of sites` = n()) |>
    ungroup() |>
    group_by(PSC) |>
    mutate(`Sites in PSC` = sum(`Number of sites`),
           Percent = `Number of sites`/`Sites in PSC`,
           Percent = label_percent(accuracy = 0.1)(Percent)
           ) |>
    ungroup()

  psc_all_phases_table
}

psc_all_phases_table(df = marthas_data_to_aggregate,
                     component = adults_patient_check_in,
                     exclude = 'No adult patients'
                     ) |>
  rename('Component 1' = adults_patient_check_in) |>
  DT::datatable(options = list(pageLength = 10))
```

#### Phase 1

**Figure 5:** Adoption stages for component 1 across phase 1 sites with adult units, by PSC

```{r, fig.height=10,fig.width=12}
likert_component(df = marthas_data_to_aggregate, 
                 component = adults_patient_check_in, 
                 exclude = 'No adult patients',
                 by_phase = 'Phase 1'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '1'),
                     component = adults_patient_check_in,
                     exclude = 'No adult patients'
                     ) |>
  rename('Component 1' = adults_patient_check_in) |>
  DT::datatable(options = list(pageLength = 10))
```

#### Phase 2

**Figure 6:** Adoption stages for component 1 across phase 2 sites with adult units, by PSC

```{r, fig.height=10,fig.width=12}
likert_component(df = marthas_data_to_aggregate, 
                 component = adults_patient_check_in, 
                 exclude = 'No adult patients',
                 by_phase = 'Phase 2'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '2'),
                     component = adults_patient_check_in,
                     exclude = 'No adult patients'
                     ) |>
  rename('Component 1' = adults_patient_check_in) |>
  DT::datatable(options = list(pageLength = 10))
```
:::

### Component 2

All staff will be able, at any time, to ask for a review from a different team if they are concerned that a patient is deteriorating, and they are not being responded to.

#### National overview

::: panel-tabset
#### Plot

**Figure 7:** Stages of adoption of component 2 across Martha’s Rule sites with adult units, by phase

```{r, fig.height=8,fig.width=10}
national_by_phase_table <- national_by_phase(df = marthas_data_to_aggregate,
                                             component = adults_independent_clinical_review,
                                             exclude = 'No adult patients'
)

plot_national_by_phase(df = national_by_phase_table, 
                       component = adults_independent_clinical_review
)
```

#### Table

**All sites**

```{r}
national_all_phases(df = marthas_data_to_aggregate, 
                    component = adults_independent_clinical_review,
                    exclude = 'No adult patients'
) |>
  kable()
```

**Phase 1 sites**

```{r}
national_all_phases(df = marthas_data_to_aggregate |> 
                      filter(Phase == 1), 
                    component = adults_independent_clinical_review,
                    exclude = 'No adult patients'
                    ) |>
  kable()
```

**Phase 2 sites**

```{r}
national_all_phases(df = marthas_data_to_aggregate |> 
                      filter(Phase == 2), 
                    component = adults_independent_clinical_review,
                    exclude = 'No adult patients'
                    ) |>
  kable()
```

#### Reports at or below stage 3

```{r}
below_stage_x(df = marthas_data_to_aggregate, 
              component = adults_independent_clinical_review,
              exclude = 'No adult patients'
              ) |>
  rename('Component 2' = adults_independent_clinical_review) |>
  kable()
```
:::

#### PSC Overview

::: panel-tabset
#### All sites

**Figure 8:** Adoption stages for component 2 across all sites with adult units, by PSC

```{r, fig.height=10,fig.width=12}
likert_component(df = marthas_data_to_aggregate, 
                 component = adults_independent_clinical_review,
                 exclude = 'No adult patients'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate,
                     component = adults_independent_clinical_review,
                     exclude = 'No adult patients'
                     ) |>
  rename('Component 2' = adults_independent_clinical_review) |>
  DT::datatable(options = list(pageLength = 10))
```

#### Phase 1

**Figure 9:** Adoption stages for component 2 across phase 1 sites with adult units, by PSC

```{r, fig.height=10,fig.width=12}
likert_component(df = marthas_data_to_aggregate, 
                 component = adults_independent_clinical_review, 
                 exclude = 'No adult patients',
                 by_phase = 'Phase 1'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '1'),
                     component = adults_independent_clinical_review,
                     exclude = 'No adult patients'
                     ) |>
  rename('Component 2' = adults_independent_clinical_review) |>
  DT::datatable(options = list(pageLength = 10))
```

#### Phase 2

**Figure 10:** Adoption stages for component 2 across phase 2 sites with adult units, by PSC

```{r, fig.height=10,fig.width=12}
likert_component(df = marthas_data_to_aggregate, 
                 component = adults_independent_clinical_review, 
                 exclude = 'No adult patients',
                 by_phase = 'Phase 2'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '2'),
                     component = adults_independent_clinical_review,
                     exclude = 'No adult patients'
                     ) |>
  rename('Component 2' = adults_independent_clinical_review) |>
  DT::datatable(options = list(pageLength = 10))
```
:::

### Component 3

This escalation route will also always be available to patients themselves, their families and carers and advertised across the hospital.

#### National overview

::: panel-tabset
#### Plot

**Figure 11:** Stages of adoption of component 3 across Martha’s Rule sites with adult units, by phase

```{r, fig.height=8,fig.width=10}
national_by_phase_table <- national_by_phase(df = marthas_data_to_aggregate,
                                             component = adults_escalation_available_to_patient_carers,
                                             exclude = 'No adult patients'
)

plot_national_by_phase(df = national_by_phase_table, 
                       component = adults_escalation_available_to_patient_carers
)
```

#### Table

**All sites**

```{r}
national_all_phases(df = marthas_data_to_aggregate, 
                    component = adults_escalation_available_to_patient_carers,
                    exclude = 'No adult patients'
) |>
  kable()
```

**Phase 1 sites**

```{r}
national_all_phases(df = marthas_data_to_aggregate |> 
                      filter(Phase == 1), 
                    component = adults_escalation_available_to_patient_carers,
                    exclude = 'No adult patients'
                    ) |>
  kable()
```

**Phase 2 sites**

```{r}
national_all_phases(df = marthas_data_to_aggregate |> 
                      filter(Phase == 2), 
                    component = adults_escalation_available_to_patient_carers,
                    exclude = 'No adult patients'
                    ) |>
  kable()
```

#### Reports at or below stage 3

```{r}
below_stage_x(df = marthas_data_to_aggregate, 
              component = adults_escalation_available_to_patient_carers,
              exclude = 'No adult patients'
              ) |>
  rename('Component 3' = adults_escalation_available_to_patient_carers) |>
  kable()
```
:::

#### PSC Overview

::: panel-tabset
#### All Sites

**Figure 12:** Adoption stages for component 3 across all sites with adult units, by PSC

```{r, fig.height=10,fig.width=12}
likert_component(df = marthas_data_to_aggregate, 
                 component = adults_escalation_available_to_patient_carers,
                 exclude = 'No adult patients'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate,
                     component = adults_escalation_available_to_patient_carers,
                     exclude = 'No adult patients'
                     ) |>
  rename('Component 3' = adults_escalation_available_to_patient_carers) |>
  DT::datatable(options = list(pageLength = 10))
```

#### Phase 1

**Figure 13:** Adoption stages for component 3 across phase 1 sites with adult units, by PSC

```{r, fig.height=10,fig.width=12}
likert_component(df = marthas_data_to_aggregate, 
                 component = adults_escalation_available_to_patient_carers, 
                 exclude = 'No adult patients',
                 by_phase = 'Phase 1'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '1'),
                     component = adults_escalation_available_to_patient_carers,
                     exclude = 'No adult patients'
                     ) |>
  rename('Component 3' = adults_escalation_available_to_patient_carers) |>
  DT::datatable(options = list(pageLength = 10))
```

#### Phase 2

**Figure 14:** Adoption stages for component 3 across phase 2 sites with adult units, by PSC

```{r, fig.height=10,fig.width=12}
likert_component(df = marthas_data_to_aggregate, 
                 component = adults_escalation_available_to_patient_carers, 
                 exclude = 'No adult patients',
                 by_phase = 'Phase 2'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '2'),
                     component = adults_escalation_available_to_patient_carers,
                     exclude = 'No adult patients'
                     ) |>
  rename('Component 3' = adults_escalation_available_to_patient_carers) |>
  DT::datatable(options = list(pageLength = 10))
```
:::

### All adult sites at stage 6 or above

```{r}
adults_stage6_plus <- marthas_data_to_aggregate |>
  select(PSC, name_trust_code, 
         name_site_code, Phase,
         starts_with('adults')
         ) |>
  filter(
    if_all(starts_with('adults'), 
           ~ str_detect(., '(?i)stage (6|7)'))
    ) 

#counts
adult_sites_at6_plus <- nrow(adults_stage6_plus)

adults_stage6_plus_table <- adults_stage6_plus |> 
  tabyl(Phase)

adults_stage6_plus_phase1 <- adults_stage6_plus_table |> 
  filter(Phase == '1') |> 
  pull(n)

adults_stage6_plus_phase2 <- adults_stage6_plus_table |> 
  filter(Phase == '2') |> 
  pull(n) 
```

This sections shows all adult sites (n = `r adult_sites_at6_plus`) reporting adoption at stage 6 or above as at `r latest_quarter`.

:::: {style="width: 45vw;"}
::: panel-tabset
#### Phase 1

There are `r adults_stage6_plus_phase1` adult sites reporting adoption at stage 6 or above.

```{r}
adults_stage6_plus |>
  filter(Phase == '1') |>
  select(PSC,
         'Trust' = name_trust_code, 
         'Site' = name_site_code,
         Phase,
         'Component 1' = adults_patient_check_in,
         'Component 2' = adults_independent_clinical_review,
         'Component 3' = adults_escalation_available_to_patient_carers) |>
  DT::datatable(options = list(pageLength = 10))
```

#### Phase 2

There are `r adults_stage6_plus_phase2` adult sites reporting adoption at stage 6 or above.

```{r}
adults_stage6_plus |>
  filter(Phase == '2') |>
  select(PSC,
         'Trust' = name_trust_code, 
         'Site' = name_site_code,
         Phase,
         'Component 1' = adults_patient_check_in,
         'Component 2' = adults_independent_clinical_review,
         'Component 3' = adults_escalation_available_to_patient_carers) |>
  DT::datatable(options = list(pageLength = 10))
```
:::
::::

## Stages of adoption across paediatric units

```{r}
paediatric_comp_1 <- national_all_phases(df = marthas_data_to_aggregate, 
                    component = paediatric_patient_check_in,
                    exclude = 'No paediatric patients'
)

total_sites_paed_units <- paediatric_comp_1 |>
  filter(Stage == 'Total') |>
  pull(`Number of sites`)
```

**Note:** The graphs and tables presented below include only sites that have paediatric units. Therefore, the total presented for this section (n = `r total_sites_paed_units`) is smaller than the total number of all Martha's Rule sites (n = `r total_sites`)

### Component 1

Patients will be asked, at least daily, about how they are feeling, and if they are getting better or worse, and this information will be acted on in a structured way.

#### National overview

::: panel-tabset
#### Plot

**Figure 15:** Stages of adoption of component 1 across Martha’s Rule sites with paediatric units, by phase

```{r, fig.height=8,fig.width=10}
national_by_phase_table <- national_by_phase(
  df = marthas_data_to_aggregate,
  component = paediatric_patient_check_in,
  exclude = 'No paediatric patients'
)

plot_national_by_phase(df = national_by_phase_table, 
                       component = paediatric_patient_check_in
)
```

#### Table

**All sites**

```{r}
paediatric_comp_1 |>
  kable()
```

**Phase 1 sites**

```{r}
national_all_phases(df = marthas_data_to_aggregate |> 
                      filter(Phase == 1), 
                    component = paediatric_patient_check_in,
                    exclude = 'No paediatric patients'
                    ) |>
  kable()
```

**Phase 2 sites**

```{r}
national_all_phases(df = marthas_data_to_aggregate |> 
                      filter(Phase == 2), 
                    component = paediatric_patient_check_in,
                    exclude = 'No paediatric patients'
                    ) |>
  kable()
```

#### Reports at or below stage 3

```{r}
below_stage_x(df = marthas_data_to_aggregate, 
              component = paediatric_patient_check_in,
              exclude = 'No paediatric patients'
              ) |>
  rename('Component 1' = paediatric_patient_check_in) |>
  kable()
```
:::

#### PSC Overview

::: panel-tabset
#### All sites

**Figure 16:** Adoption stages for component 1 across all sites with paediatric units, by PSC

```{r, fig.height=10,fig.width=12}
likert_component(df = marthas_data_to_aggregate, 
                 component = paediatric_patient_check_in, 
                 exclude = 'No paediatric patients'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate,
                     component = paediatric_patient_check_in,
                     exclude = 'No paediatric patients'
                     ) |>
  rename('Component 1' = paediatric_patient_check_in) |>
  DT::datatable(options = list(pageLength = 10))
```

#### Phase 1

**Figure 17:** Adoption stages for component 1 across phase 1 sites with paediatric units, by PSC

```{r, fig.height=10,fig.width=12}
likert_component(df = marthas_data_to_aggregate, 
                 component = paediatric_patient_check_in, 
                 exclude = 'No paediatric patients',
                 by_phase = 'Phase 1'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '1'),
                     component = paediatric_patient_check_in,
                     exclude = 'No paediatric patients'
                     ) |>
  rename('Component 1' = paediatric_patient_check_in) |>
  DT::datatable(options = list(pageLength = 10))
```

#### Phase 2

**Figure 18:** Adoption stages for component 1 across phase 2 sites with paediatric units, by PSC

```{r, fig.height=10,fig.width=12}
likert_component(df = marthas_data_to_aggregate, 
                 component = paediatric_patient_check_in, 
                 exclude = 'No paediatric patients',
                 by_phase = 'Phase 2'
                   )

```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '2'),
                     component = paediatric_patient_check_in,
                     exclude = 'No paediatric patients'
                     ) |>
  rename('Component 1' = paediatric_patient_check_in) |>
  DT::datatable(options = list(pageLength = 10))
```
:::

### Component 2

All staff will be able, at any time, to ask for a review from a different team if they are concerned that a patient is deteriorating, and they are not being responded to.

#### National overview

::: panel-tabset
#### Plot

**Figure 19:** Stages of adoption of component 2 across Martha’s Rule sites with paediatric units, by phase

```{r, fig.height=8,fig.width=10}
national_by_phase_table <- national_by_phase(df = marthas_data_to_aggregate,
                                             component = paediatric_independent_clinical_review,
                                              exclude = 'No paediatric patients'
)

plot_national_by_phase(df = national_by_phase_table, 
                       component = paediatric_independent_clinical_review
)
```

#### Table

**All sites**

```{r}
national_all_phases(df = marthas_data_to_aggregate, 
                    component = paediatric_independent_clinical_review,
                    exclude = 'No paediatric patients'
) |>
  kable()
```

**Phase 1 sites**

```{r}
national_all_phases(df = marthas_data_to_aggregate |> 
                      filter(Phase == 1), 
                    component = paediatric_independent_clinical_review,
                    exclude = 'No paediatric patients'
                    ) |>
  kable()
```

**Phase 2 sites**

```{r}
national_all_phases(df = marthas_data_to_aggregate |> 
                      filter(Phase == 2), 
                    component = paediatric_independent_clinical_review,
                    exclude = 'No paediatric patients'
                    ) |>
  kable()
```

#### Reports at or below stage 3

```{r}
below_stage_x(df = marthas_data_to_aggregate, 
              component = paediatric_independent_clinical_review,
              exclude = 'No paediatric patients'
              ) |>
  rename('Component 2' = paediatric_independent_clinical_review) |>
  kable()
```
:::

#### PSC Overview

::: panel-tabset
#### All sites

**Figure 20:** Adoption stages for component 2 across all sites with paediatric units, by PSC

```{r, fig.height=10,fig.width=12}
likert_component(df = marthas_data_to_aggregate, 
                 component = paediatric_independent_clinical_review, 
                 exclude = 'No paediatric patients'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate,
                     component = paediatric_independent_clinical_review,
                     exclude = 'No paediatric patients'
                     ) |>
  rename('Component 2' = paediatric_independent_clinical_review) |>
  DT::datatable(options = list(pageLength = 10))
```

#### Phase 1

**Figure 21:** Adoption stages for component 2 across phase 1 sites with paediatric units, by PSC

```{r, fig.height=10,fig.width=12}
likert_component(df = marthas_data_to_aggregate, 
                 component = paediatric_independent_clinical_review, 
                 exclude = 'No paediatric patients',
                 by_phase = 'Phase 1'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '1'),
                     component = paediatric_independent_clinical_review,
                     exclude = 'No paediatric patients'
                     ) |>
  rename('Component 2' = paediatric_independent_clinical_review) |>
  DT::datatable(options = list(pageLength = 10))
```

#### Phase 2

**Figure 22:** Adoption stages for component 2 across phase 2 sites with paediatric units, by PSC

```{r, fig.height=10,fig.width=12}
likert_component(df = marthas_data_to_aggregate, 
                 component = paediatric_independent_clinical_review, 
                 exclude = 'No paediatric patients',
                 by_phase = 'Phase 2'
                   )

```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '2'),
                     component = paediatric_independent_clinical_review,
                     exclude = 'No paediatric patients'
                     ) |>
  rename('Component 2' = paediatric_independent_clinical_review) |>
  DT::datatable(options = list(pageLength = 10))
```
:::

### Component 3

This escalation route will also always be available to patients themselves, their families and carers and advertised across the hospital.

#### National overview

::: panel-tabset
#### Plot

**Figure 23:** Stages of adoption of component 3 across Martha’s Rule sites with paediatric units, by phase

```{r, fig.height=8,fig.width=10}
national_by_phase_table <- national_by_phase(df = marthas_data_to_aggregate,
                                             component = paediatric_escalation_available_to_patient_carers,
                                              exclude = 'No paediatric patients'
)

plot_national_by_phase(df = national_by_phase_table, 
                       component = paediatric_escalation_available_to_patient_carers
)
```

#### Table

**All sites**

```{r}
national_all_phases(df = marthas_data_to_aggregate, 
                    component = paediatric_escalation_available_to_patient_carers,
                    exclude = 'No paediatric patients'
) |>
  kable()
```

**Phase 1 sites**

```{r}
national_all_phases(df = marthas_data_to_aggregate |> 
                      filter(Phase == 1), 
                    component = paediatric_escalation_available_to_patient_carers,
                    exclude = 'No paediatric patients'
                    ) |>
  kable()
```

**Phase 2 sites**

```{r}
national_all_phases(df = marthas_data_to_aggregate |> 
                      filter(Phase == 2), 
                    component = paediatric_escalation_available_to_patient_carers,
                    exclude = 'No paediatric patients'
                    ) |>
  kable()
```

#### Reports at or below stage 3

```{r}
below_stage_x(df = marthas_data_to_aggregate, 
              component = paediatric_escalation_available_to_patient_carers,
              exclude = 'No paediatric patients'
              ) |>
  rename('Component 3' = paediatric_escalation_available_to_patient_carers) |>
  kable()
```
:::

#### PSC Overview

::: panel-tabset
#### All sites

**Figure 24:** Adoption stages for component 3 across all sites with paediatric units, by PSC

```{r, fig.height=10,fig.width=12}
likert_component(df = marthas_data_to_aggregate, 
                 component = paediatric_escalation_available_to_patient_carers, 
                 exclude = 'No paediatric patients'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate,
                     component = paediatric_escalation_available_to_patient_carers,
                     exclude = 'No paediatric patients'
                     ) |>
  rename('Component 3' = paediatric_escalation_available_to_patient_carers) |>
  DT::datatable(options = list(pageLength = 10))
```

#### Phase 1

**Figure 25:** Adoption stages for component 3 across phase 1 sites with paediatric units, by PSC

```{r, fig.height=10,fig.width=12}
likert_component(df = marthas_data_to_aggregate, 
                 component = paediatric_escalation_available_to_patient_carers, 
                 exclude = 'No paediatric patients',
                 by_phase = 'Phase 1'
                   )
```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '1'),
                     component = paediatric_escalation_available_to_patient_carers,
                     exclude = 'No paediatric patients'
                     ) |>
  rename('Component 3' = paediatric_escalation_available_to_patient_carers) |>
  DT::datatable(options = list(pageLength = 10))
```

#### Phase 2

**Figure 26:** Adoption stages for component 3 across phase 2 sites with paediatric units, by PSC

```{r, fig.height=10,fig.width=12}
likert_component(df = marthas_data_to_aggregate, 
                 component = paediatric_escalation_available_to_patient_carers, 
                 exclude = 'No paediatric patients',
                 by_phase = 'Phase 2'
                   )

```

**Underlying data**

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate |> filter(Phase == '2'),
                     component = paediatric_escalation_available_to_patient_carers,
                     exclude = 'No paediatric patients'
                     ) |>
  rename('Component 3' = paediatric_escalation_available_to_patient_carers) |>
  DT::datatable(options = list(pageLength = 10))
```
:::

### All paediatric sites at stage 6 or above

```{r}
paeds_stage6_plus <- marthas_data_to_aggregate |>
  select(PSC, name_trust_code, 
         name_site_code, Phase,
         starts_with('paed')
         ) |>
  filter(
    if_all(starts_with('paed'), 
           ~ str_detect(., '(?i)stage (6|7)'))
    ) 

#counts
paeds_sites_at6_plus <- nrow(paeds_stage6_plus)

paeds_stage6_plus_table <- paeds_stage6_plus |> 
  tabyl(Phase)

paeds_stage6_plus_phase1 <- paeds_stage6_plus_table |> 
  filter(Phase == '1') |> 
  pull(n)

paeds_stage6_plus_phase2 <- paeds_stage6_plus_table |> 
  filter(Phase == '2') |> 
  pull(n) 
```

This sections shows all paediatric sites (n = `r paeds_sites_at6_plus`) reporting adoption at stage 6 or above as at `r latest_quarter`.

:::: {style="width: 45vw;"}
::: panel-tabset
#### Phase 1

There are `r paeds_stage6_plus_phase1` paediatric sites reporting adoption at stage 6 or above.

```{r}
paeds_stage6_plus |>
  filter(Phase == '1') |>
  select(PSC,
         'Trust' = name_trust_code, 
         'Site' = name_site_code,
         Phase,
         'Component 1' = paediatric_patient_check_in,
         'Component 2' = paediatric_independent_clinical_review,
         'Component 3' = paediatric_escalation_available_to_patient_carers) |>
  DT::datatable(options = list(pageLength = 10))
```

#### Phase 2

There are `r paeds_stage6_plus_phase2` paediatric sites reporting adoption at stage 6 or above.

```{r}
paeds_stage6_plus |>
  filter(Phase == '2') |>
  select(PSC,
         'Trust' = name_trust_code, 
         'Site' = name_site_code,
         Phase,
         'Component 1' = paediatric_patient_check_in,
         'Component 2' = paediatric_independent_clinical_review,
         'Component 3' = paediatric_escalation_available_to_patient_carers) |>
  DT::datatable(options = list(pageLength = 10))
```
:::
::::

## All adult data

::: {style="width: 55vw;"}
```{r}
# stages to colour
stages <- c('Stage 0 - No contact', 'Stage 1 - Knowledge', 'Stage 2 - Interest')

marthas_data_to_aggregate |>
  select(PSC, name_trust_code, name_site_code, 
         Phase, matches('adults')) |>
  rename('Trust' = name_trust_code, 
         'Site' = name_site_code,
         'Component 1' = adults_patient_check_in,
         'Component 2' = adults_independent_clinical_review,
         'Component 3' = adults_escalation_available_to_patient_carers
         ) |>
    DT::datatable(options = 
                    list(pageLength = 10,
                         columnDefs = list(
                           list(width = '110px', targets = c(2, 3)),
                           list(width = '50px', targets = c(4)),
                           list(width = '80px', targets = c(5, 6, 7))
                           )
                         )
                  ) |>
    DT::formatStyle(
      columns = c('Component 1', 'Component 2','Component 3'),
      backgroundColor = styleEqual(
        levels = stages,
        c('#D73027', '#F46D43', '#FDAE61')
      )
    ) 
```
:::

::: {style="width: 55vw;"}
## All paediatric data

```{r}
marthas_data_to_aggregate |>
  select(PSC, name_trust_code, name_site_code, Phase, matches('paediatric')) |> 
  rename('Trust' = name_trust_code, 
         'Site' = name_site_code,
         'Component 1' = paediatric_patient_check_in,
         'Component 2' = paediatric_independent_clinical_review,
         'Component 3' = paediatric_escalation_available_to_patient_carers
         ) |>
    DT::datatable(options = 
                    list(pageLength = 10,
                         columnDefs = list(
                           list(width = '110px', targets = c(2, 3)),
                           list(width = '50px', targets = c(4)),
                           list(width = '80px', targets = c(5, 6, 7))
                           )
                         )
                  ) |>
  DT::formatStyle(
    columns = c('Component 1', 'Component 2','Component 3'),
    backgroundColor = styleEqual(
      levels = stages,
      c('#D73027', '#F46D43', '#FDAE61')
    )
  )
```
:::
