---
title: "NatPatSIP - Martha's Rule QART Report"
author: "Patient Safety Analysis Team"
format: 
   html:  
     embed-resources: true
     df-print: kable
     toc: true
     toc-location: left
     number-sections: true
execute:
  echo: false
  warning: false
params: 
  reporting_quarter: '2024/25 Q6'
  master_files_folder: '1. Master files'
  quarter_string: '25_26_Q1'
  marthas_data_file: 'marthas_psc_submissions_2025_26_Q1_processed_2025_09_08_184653.csv'
---


```{r}
# read in required packages
library(here)
library(tidyverse)
library(janitor)
library(lubridate)
library(ggplot2)
library(readxl)
library(forcats)
library(scales)
library(ggsci)
library(viridis)
library(Microsoft365R)
library(ggplot2) # likert plots
library(ggstats)# likert plots
library(kableExtra) # tables 
library(DT) # sortable tables

source('config_sharepoint_location.R')


# params
reporting_quarter <- params$reporting_quarter
master_files_folder <- params$master_files_folder
quarter_string <- params$quarter_string
marthas_data_file <- params$marthas_data_file

# bring site look up with postcode info and get geo inpho
site_postcode_list <- read.csv(here('lookups', 'psc_trust_site_lookup.csv')) |>
  select(trust_site_code, api_site_postcode) 

library(PostcodesioR)

postcodes_lookup_table <- lapply(unique(site_postcode_list$api_site_postcode),
                                 PostcodesioR::postcode_lookup) |>
  bind_rows() |>
  right_join(site_postcode_list, by = c('postcode' = 'api_site_postcode'))
```


```{r}
# NOTE block below is a place holder that will coem in place in 2526 Q2
# go into share point and download latest data
#master_files_dr <- chosenlib$get_item(glue::glue("{base_url}/{master_files_folder}"))

# marthas_data_file <- master_files_dr$list_files() |> 
#   select(name) |> 
#   filter(str_detect(name, 
#                     # below parameter is generated in read_qart_submissions.R
#                     quarter_string)
#          ) |>
#   as.vector() |>
#   unlist() |> 
#   unname()
# 
# marthas_identified_data <- master_files_dr$get_item(str_glue({marthas_data_file}))
# 
# marthas_identified_data$download(dest = here("data", str_glue({marthas_data_file})), 
#                        overwrite = T)

# marthas_data <- read.csv(here(str_glue('data/{marthas_data_file}'))) |>
#   mutate(updated_psc_name = str_remove_all(updated_psc_name, 'HIN'))

marthas_data_raw <- read.csv(here(str_glue('output/{marthas_data_file}'))) |>
  mutate(updated_psc_name = str_remove_all(updated_psc_name, ' HIN')) 

marthas_data_postcode  <- marthas_data_raw |>
  left_join(postcodes_lookup_table, by = c('site_code' = 'trust_site_code')) |>
  relocate(postcode, .after = name_of_site)
```

## Note

The visualisations and breakdowns below are experimental and for internal use only. Data presented is for `r reporting_quarter`.

## Distribution of sites by phase 

Missing data for two phase 2 sites here: 

- RGD03 - Leeds General Infirmary
- RAN01 - The Royal National Orthopaedic Hospital (Stanmore)

```{r}
# free styling aggregations and visualisations
marthas_data_to_aggregate <- marthas_data_raw |>
  rename('PSC' = updated_psc_name,
         'Phase' = phase)

marthas_data_to_aggregate |>
  tabyl(Phase, sort = T) |> 
  rename('Count' = n, 'Percent' = percent) |> 
  mutate(Percent = label_percent(accuracy = 0.1)(Percent)) |>
  kable(align = 'lll')
```



```{r}
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
# Get United Kingdom boundaries at 1:10m scale
uk <- ne_states(country = "United Kingdom", returnclass = "sf")

# Filter for England
england <- uk %>% filter(geonunit == "England")

# outilinee of england
england_outline <- england %>%
  summarise(geometry = st_union(geometry))

# make data sf
trust_sites_sf <- sf::st_as_sf(marthas_data_postcode , coords = c('longitude', 'latitude'),
                      crs = 4326)
```

**Figure :** Martha's Rule sites in England, by phase

```{r, fig.height=8,fig.width=8}
ggplot() +
  geom_sf(data = england_outline, fill = "white", color = "black") +
  geom_sf(data = trust_sites_sf, 
          aes(colour = factor(phase)),
          size = 2,
          alpha = 0.7
          ) +
  scale_colour_manual(values = c('#801650', '#28A197'),
                      labels = c("Phase 1", "Phase 2")  # Custom legend labels
                    ) +
  theme_minimal()  +
  theme(axis.text = element_blank(),
        panel.grid = element_blank(),
        legend.title = element_blank()
        )
```

## Share of sites by PSC

::: {.panel-tabset}

```{r}
# maybe make this into a bar plot, 2 shades of clue for each phase that make up the totality of sites 
share_of_sites <- marthas_data_to_aggregate |>
  tabyl(PSC, Phase) |>
  rename('Phase 1' = `1`,
         'Phase 2' = `2`
         ) 

share_of_sites_long <- share_of_sites |>
  pivot_longer(cols = starts_with('Phase'), 
               names_to = 'Phase', 
               values_to = 'Number of sites')

#share of site table 
share_of_sites_table <- share_of_sites |>
  adorn_totals('both') |>
  mutate(Percent = label_percent(accuracy = 0.1)(Total/210))
```

### Plot

**Figure :** Breakdown of Martha's rule sites by PSC and phase

```{r, fig.height=10,fig.width=10}
share_of_sites_for_plot <- share_of_sites_long |>
  left_join(share_of_sites_table |> select(PSC, Percent), by = 'PSC')

share_of_sites_perc_labels <- share_of_sites_for_plot |>
  group_by(PSC) |> 
  summarise('Number of sites' = sum(`Number of sites`), Percent = first(Percent)) |>
  ungroup()

ggplot(data = share_of_sites_for_plot,
       aes(x = `Number of sites`, y = forcats::fct_rev(PSC)
           )
       ) +
  geom_bar(aes( fill = forcats::fct_rev(Phase)),stat = 'identity', alpha = 0.7) +
  geom_text(data = share_of_sites_perc_labels,
            aes(x = `Number of sites`, 
                y = forcats::fct_rev(PSC), 
                label = Percent
                ),
            hjust = -0.1,
            position=position_dodge2(width=0.9),
            size = 4.5
            ) +
  scale_x_continuous(expand = expansion(mult = c(0.02,0.11))) +
  scale_fill_manual(values = c('#28A197', '#801650')
                    ) +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(color = "black", size = 13),
        axis.text.x = element_text(color = "black", size = 13),
        axis.text.y = element_text(color = "black", size = 13),
        legend.text = element_text(color = "black", size = 13),
        panel.grid.minor.x = element_blank(),
        axis.ticks = element_line(color = "grey"),
        legend.title = element_blank())
```

### Table

```{r}
share_of_sites_table |> 
  kable(align = 'llll')
```

:::

## Stages of adoption across adult units 

### Component 1

Patients will be asked, at least daily, about how they are feeling, and if they are getting better or worse, and this information will be acted on in a structured way.

#### National overview

Lot of the data in the middle stages for either phase, even though sites in phase 1 are assumed to have been in the programme longer (check this assumption)

::: {.panel-tabset}

#### Plot

**Figure :** Stages of adoption of component 1 (in adults) across Martha's rule sites, by phase

```{r}
# function to aggregate by phase adn component
national_by_phase <- function(df, component, exclude){
  # quote variable
  component <- enquo(component)
  
  national_by_phase <- df |>
    mutate(Phase = as.character(Phase)) |>
    # exclude these from percetnage calcualtion
    filter(!!component != exclude
           ) |>
    group_by(Phase, !!component) |>
    summarise(number_of_sites = n()) |>
    group_by(Phase) |>
    mutate(group_size_by_phase = sum(number_of_sites)) |>
    arrange(!!component )
  
  national_by_phase
}

# function to plot aggregated table from national_by_phase()
plot_national_by_phase <- function(df, component){
  
  # quote variable
  component <- enquo(component)
  
  plot_national_by_phase <- ggplot(
    data = df,
    aes(
      y = number_of_sites,
      x = factor(!!component)
      )
    ) +
  geom_bar(aes(fill = fct_rev(factor(Phase))),
           stat = 'identity',
           alpha = 0.7,
           position = position_dodge()) +
  labs(x = "Stage",
       y = "Number of sites"
       )+
  scale_fill_manual(values = c('#28A197', '#801650'),
                    name = "Phase",  # Custom legend title
                    labels = c("Phase 2", "Phase 1")  # Custom legend labels
                    )+
  scale_x_discrete(labels = label_wrap(25)) +
  theme_minimal() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_text(color = "black", size = 13),
        axis.text.x = element_text(color = "black", size = 13,
                                   angle = 90),
        axis.text.y = element_text(color = "black", size = 13),
        panel.grid.minor.x = element_blank(),
        axis.ticks = element_line(color = "grey"),
        legend.title = element_blank(),
        legend.text = element_text(color = "black", size = 13)
        )
  
  plot_national_by_phase
}
```

```{r, fig.height=6,fig.width=10}
# deploy functions
national_by_phase_table <- national_by_phase(df = marthas_data_to_aggregate,
                                             component = adults_patient_check_in,
                                              exclude = 'No adult patients'
                                             )

plot_national_by_phase(df = national_by_phase_table, 
                       component = adults_patient_check_in
                         )
```

#### Table 

Note: fewer sites here (203) than overall, given there are sites with no adult units 

```{r}
national_all_phases <- function(df, component, exclude){
  # quote
  component <- enquo(component)
  
  national_all_phases <- df |>
    # exclude these from percetnage calcualtion
    filter(!!component != exclude) |>
    tabyl(!!component) |>
    rename('Number of sites' = n, 'Percent' = percent, 'Stage' = !!component) |> 
    adorn_totals('row') |>
    mutate(Percent = label_percent(accuracy = 0.1)(Percent)) 
  
  national_all_phases
}

national_all_phases(df = marthas_data_to_aggregate, 
                    component = adults_patient_check_in,
                    exclude = 'No adult patients'
                    ) |>
  kable()
```

#### Reports at or below stage 2
```{r}
below_stage_x <- function(df, component, exclude, stage_filter = '(?i)stage (0|1|2)'){
  
  component <- enquo(component)
  
  below_stage_x <- df |>
    mutate(Phase = as.character(Phase)) |>
    filter(!!component != exclude,
           str_detect(adults_patient_check_in, stage_filter)) |>
    select(PSC, site_code, name_of_site, Phase, !!component) |>
    arrange(!!component , Phase, PSC)
    
  below_stage_x
    
}

below_stage_x(df = marthas_data_to_aggregate, 
              component = adults_patient_check_in,
              exclude = 'No adult patients'
              ) |>
  kable()
```

:::

#### PSC Overview

::: {.panel-tabset}

#### Plot

**Figure :** Breakdown of adoption states for component 1 in adults, by PSC

```{r, fig.height=9,fig.width=11}
likert_component <- function(df, component, exclude){
  
  component <- enquo(component)
  
  psc_all_phases <- df |>
    # exclude these from percetnage calcualtion
    filter(!!component != exclude) |>
    select(PSC, site_code, !!component) |>
    # note change from quoted component to local component
    mutate(component = factor(!!component))

  psc_all_phases_likert <- psc_all_phases |>
    select(-!!component) |>
    pivot_wider(names_from = PSC,
                values_from = component)
  
  plot <- gglikert(psc_all_phases_likert |> 
             # drop so we can graph by PSC correctly
             select(-site_code),
             # exclude second set of percentages to the sides
             add_totals = F,
           ) +
    scale_fill_brewer(palette = "RdYlGn") +
    guides(fill = guide_legend(nrow = 3, byrow = T))+
    theme(axis.text.x = element_blank(),
          axis.text.y = element_text(color = "black", size = 13),
          legend.text = element_text(color = "black", size = 13)
            )
  
  plot
}

likert_component(df = marthas_data_to_aggregate, 
                 component = adults_patient_check_in, 
                 exclude = 'No adult patients'
                   )
```

#### Table

```{r}
psc_all_phases_table <- function(df, component, exclude){
  component <- enquo(component)

  psc_all_phases_table <- df |>
    # exclude these from percetnage calcualtion
    filter(!!component != exclude) |>
    group_by(PSC, !!component, .drop = F) |>
    summarise(n_sites = n()) |>
    ungroup() |>
    group_by(PSC) |>
    mutate(total_n_psc_sites = sum(n_sites),
           Percent = n_sites/total_n_psc_sites,
           Percent = label_percent(accuracy = 0.1)(Percent)
           ) |>
    ungroup()

  psc_all_phases_table
}

psc_all_phases_table(df = marthas_data_to_aggregate,
                     component = adults_patient_check_in,
                     exclude = 'No adult patients'
                     ) |>
  DT::datatable(options = list(pageLength = 10))
```

:::  

## Stages of adoption across paediatric units 

### Component 1

Patients will be asked, at least daily, about how they are feeling, and if they are getting better or worse, and this information will be acted on in a structured way.

#### National overview

TEXT

::: {.panel-tabset}

#### Plot

**Figure :** Stages of adoption of component 1 (in paediatric) across Martha's rule sites, by phase

```{r, fig.height=6,fig.width=10}
national_by_phase_table <- national_by_phase(df = marthas_data_to_aggregate,
                                             component = paediatric_patient_check_in,
                                              exclude = 'No paediatric patients'
)

plot_national_by_phase(df = national_by_phase_table, 
                       component = paediatric_patient_check_in
)
```


#### Table 

Note: fewer sites here (180) than overall, given there are sites with no paediatric units 

```{r}
national_all_phases(df = marthas_data_to_aggregate, 
                    component = paediatric_patient_check_in,
                    exclude = 'No paediatric patients'
) |>
  kable()
```


#### Reports at or below stage 2

```{r}
below_stage_x(df = marthas_data_to_aggregate, 
              component = paediatric_patient_check_in,
              exclude = 'No paediatric patients'
) |>
  kable()
```

:::

#### PSC Overview

::: {.panel-tabset}

#### Plot

**Figure :** Breakdown of adoption states for component 1 in adults, by PSC

```{r, fig.height=9,fig.width=11}
likert_component(df = marthas_data_to_aggregate, 
                 component = paediatric_patient_check_in, 
                 exclude = 'No paediatric patients'
                   )
```

#### Table

```{r}
psc_all_phases_table(df = marthas_data_to_aggregate,
                     component = adults_patient_check_in,
                     exclude = 'No paediatric patients'
                     ) |>
  DT::datatable(options = list(pageLength = 10))
```

:::

## All adult data 

<div style="width: 80vw;">

```{r}
# not sure this is helpful
stages <- c('Stage 0 - No contact', 'Stage 1 - Knowledge', 'Stage 2 - Interest')

marthas_data_name_site_consolidated <- marthas_data_to_aggregate |>
  # consolidate org code and names for table printing
  mutate(name_trust_code = paste0(name_of_trust, ' (', trust_code, ')'),
         name_site_code = paste0(name_of_site, ' (', site_code, ')'),
         .before = Phase) |>
  select(-c(trust_code, name_of_trust, site_code, name_of_site))

marthas_data_name_site_consolidated |>
  select(PSC, name_trust_code, name_site_code, Phase, matches('adults')) |>
    DT::datatable(options = list(pageLength = 21)) |>
    DT::formatStyle(
      columns = c('adults_patient_check_in', 'adults_independent_clinical_review',
                  'adults_escalation_available_to_patient_carers'),
      backgroundColor = styleEqual(
        levels = stages,
        c('red', 'tomato', 'orange')
      )
    )
```

</div>

<div style="width: 75vw;">

## All paediatric data 
```{r}
marthas_data_name_site_consolidated |>
  select(PSC, name_trust_code, name_site_code, Phase, matches('paediatric')) |> 
  DT::datatable(options = list(pageLength = 21)) |>
  DT::formatStyle(
    columns = c('paediatric_patient_check_in', 'paediatric_independent_clinical_review',
                'paediatric_escalation_available_to_patient_carers'),
    backgroundColor = styleEqual(
      levels = stages,
      c('red', 'tomato', 'orange')
    )
  )
```

</div>
